Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-06-19 doc/network.tex
abstract?

@@

An epidemic can be characterized by its speed (exponential growth rate, $r$) and its strength, characterized by its reproductive number, \RR.
The reproductive number, defined as the average number of secondary cases arising from a primary case, is of particular interest as it provides information about the final size of an epidemic \citep{anderson1991infectious, diekmann1990definition}.
\jd{Took out a comment about tracing and tried to clarify.}
However, estimating the reproductive number directly from disease natural history requires detailed knowledge which is not often available, particularly early in an outbreak \citep{dietz1993estimation}.
Instead, the reproductive number can be \emph{indirectly} estimated from the exponential growth rate, which can be estimated from incidence data \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}.
These two quantities are linked by generation-interval distributions \citep{wearing2005appropriate, svensson2007note, roberts2007model, wallinga2007generation, park2019practical}.

>>>>

\section*{Abstract}

Generation intervals, defined as the time between when an individual is infected and when that individual infects another person, provides a link between two key quantities that describe an epidemic: the reproductive number, $\RR$, and the rate of exponential growth, $r$.
Often, generation intervals are measured through contact tracing by identifying who infected whom;
we study how these observed intervals vary both spatially and temporally.
Temporal variation in generation intervals is caused by the censoring in the observation process through contact tracing, whereas spatial variation in generation intervals reflects the patterns of spread of disease.
Therefore, correcting for temporal variation, but not for spatial variation, provides an ``effective'' generation-interval distribution, which correctly links $r$ and $\RR$.

@@

An epidemic can be characterized by the exponential growth rate, $r$, and the reproductive number, \RR.
The reproductive number, defined as the average number of secondary cases arising from a primary case, is of particular interest as it provides information about the final size of an epidemic \citep{anderson1991infectious, diekmann1990definition}.
\jd{Took out a comment about tracing and tried to clarify.}
However, estimating the reproductive number directly from disease natural history requires detailed knowledge which is not often available, particularly early in an outbreak \citep{dietz1993estimation}.
Instead, the reproductive number can be \emph{indirectly} estimated from the exponential growth rate, which can be estimated from incidence data \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}.
These two quantities are linked by generation-interval distributions \citep{wearing2005appropriate, svensson2007note, roberts2007model, wallinga2007generation, park2019practical}.
